---
title: "Fire Incident Dispatch Data Analysis"
author: "Zuliani Riccardo"
date: "12/12/2023"
output: 
  html_document: 
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE) #, cache=TRUE
setwd("C:/Users/ricca/Desktop/UNI/Magistrale/Anno3/Statistical_Inference_and_Learning/SIL Projcet/Statistical_Inference_Learning_Project/Fire_Incident_Dispatch_Analysis")
```

# Loads & Install Packages

```{r}
if (!require("nnet")) install.packages("nnet")
if (!require("summarytools")) install.packages("summarytools")
if (!require("dplyr")) install.packages("dplyr")
if (!require("ggplot2")) install.packages("ggplot2")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("lubridate")) install.packages("lubridate")
if (!require("mapview")) install.packages("mapview")
if (!require("sf")) install.packages("sf")
if (!require("geojsonio")) install.packages("geojsonio")
if (!require("leaflet")) install.packages("leaflet")
if (!require("broom")) install.packages("broom")
if (!require("plotly")) install.packages("plotly")

library(nnet)
library(summarytools)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(mapview)
library(sf)
library(geojsonio)
library(leaflet) 
library(broom)
library(plotly)
```


# Dataset description

The Fire Incident Dispatch Data file contains data that is generated by the Starfire Computer Aided Dispatch System. The data spans from the time the incident is created in the system to the time the incident is closed in the system. It covers information about the incident as it relates to the assignment of resources and the Fire Department’s response to the emergency. To protect personal identifying information in accordance with the Health Insurance Portability and Accountability Act (HIPAA), specific locations of incidents are not included and have been aggregated to a higher level of detail.

In this analysis we have restricted the analysis only on the *last 50.000 observations*.

1. **STARFIRE_INCIDENT_ID**: An incident identifier comprising the 5 character julian date, 4 character alarm box number, 2 character number of incidents at the box so far for the day, 1 character borough code , 4 character sequence number. 
2. **INCIDENT_DATETIME**: The date and time of the incident.
3. **ALARM_BOX_BOROUGH**: The borough of the alarm box.
4. **ALARM_BOX_LOCATION**: The location of the alarm box.
5. **ALARM_BOX**: The alarm box number.
6. **INCIDENT_BOROUGH**: The borough of the incident.
7. **ZIPCODE**: The zip code of the incident.
8. **POLICEPRECINCT**: The police precinct of the incident.
9. **CITYCOUNCILDISTRICT**: The city council district.
10. **COMMUNITYDISTRICT**: The community district.
11. **COMMUNITYSCHOOLDISTRICT**: The community school district.
12. **CONGRESSIONALDISTRICT**: The congressional district.
13. **ALARM_SOURCE_DESCRIPTION_TX**: The description of the alarm source.
14. **ALARM_LEVEL_INDEX_DESCRIPTION**: The alarm level index.
15. **HIGHEST_ALARM_LEVEL**: The highest alarm level.
16. **INCIDENT_CLASSIFICATION**: The incident classification.
17. **INCIDENT_CLASSIFICATION_GROUP**: The incident classification roll up group. (*response*)
18. **DISPATCH_RESPONSE_SECONDS_QY**: The elapsed time in seconds between the incident_datetime and the first_assignment_datetime.
19. **FIRST_ASSIGNMENT_DATETIME**: The date and time of the first unit assignment.
20. **FIRST_ACTIVATION_DATETIME**: The date and time of the first unit acknowledgement of the assignment.
21. **FIRST_ON_SCENE_DATETIME**: The date and time of the first unit at the scene of the incident.
22. **INCIDENT_CLOSE_DATETIME**: The date and time that the incident was closed in the dispatch system.
23. **VALID_DISPATCH_RSPNS_TIME_INDC**: Indicates that the components comprising the generation of the DISPATCH_RESPONSE_SECONDS_QY are valid.
24. **VALID_INCIDENT_RSPNS_TIME_INDC**: Indicates that the components comprising the generation of the INCIDENT_RESPONSE_SECONDS_QY are valid.
25. **INCIDENT_RESPONSE_SECONDS_QY**: The elapsed time in seconds between the incident_datetime and the first_onscene_datetime.
26. **INCIDENT_TRAVEL_TM_SECONDS_QY**: The elapsed time in seconds between the first_assignment_datetime and the first_onscene_datetime.
27. **ENGINES_ASSIGNED_QUANTITY**: The number of engine units assigned to the incident.
28. **LADDERS_ASSIGNED_QUANTITY**: The number of ladder units assigned to the incident.
29. **OTHER_UNITS_ASSIGNED_QUANTITY**: The number of  units that are not engines or ladders that were assigned to the incident.

For this analysis we use as response the **INCIDENT_CLASSIFICATION_GROUP** which contain the main grups of incidents. They are the following one:

1. **Structural Fires**
2. **NonStructural Fires**
3. **NonMedical Emergencies**
4. **NonMedical MFAs**
5. **Medical Emergencies**
6. **Medical MFAs**


# Data Exlporation and Cleaning

The first step is always to read the dataset and plot the first 5 observations
```{r, cache=TRUE}
fire_data <- read.csv("datasets/Fire_Incident_Dispatch_Data_last_50k.csv")

head(fire_data)
```

Use `dfSummary` from summarytool in order to have a complete and clear sumamry of the dataset.
```{r}
print(dfSummary(fire_data, 
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.75),
                method = 'render')
#print(dfSummary(fire_data), method = 'render')
```

Now we rename all the columns in order to be smaller whenever we plot graphs.
```{r}
fire_data <- fire_data %>%
            rename(id = STARFIRE_INCIDENT_ID, datetime = INCIDENT_DATETIME, al_borough = ALARM_BOX_BOROUGH, al_number = ALARM_BOX_NUMBER, 
                   al_location = ALARM_BOX_LOCATION, inc_borough = INCIDENT_BOROUGH, zipcode = ZIPCODE, pol_prec = POLICEPRECINCT,
                   city_con_dist = CITYCOUNCILDISTRICT, commu_dist = COMMUNITYDISTRICT, commu_sc_dist = COMMUNITYSCHOOLDISTRICT,
                   cong_dist = CONGRESSIONALDISTRICT, al_source_desc = ALARM_SOURCE_DESCRIPTION_TX, al_index_desc = ALARM_LEVEL_INDEX_DESCRIPTION,
                   highest_al_level = HIGHEST_ALARM_LEVEL, inc_class = INCIDENT_CLASSIFICATION, inc_class_group = INCIDENT_CLASSIFICATION_GROUP,
                   disp_resp_qy = DISPATCH_RESPONSE_SECONDS_QY, first_ass_datetime = FIRST_ASSIGNMENT_DATETIME,
                   first_act_datetime = FIRST_ACTIVATION_DATETIME,  first_onscene_datetime = FIRST_ON_SCENE_DATETIME,
                   inc_close_datetime = INCIDENT_CLOSE_DATETIME, disp_resp_time_indc = VALID_DISPATCH_RSPNS_TIME_INDC,
                   inc_resp_sec_indc = VALID_INCIDENT_RSPNS_TIME_INDC, inc_resp_sec_qy = INCIDENT_RESPONSE_SECONDS_QY,
                   inc_travel_sec_qy = INCIDENT_TRAVEL_TM_SECONDS_QY, engines_assigned = ENGINES_ASSIGNED_QUANTITY,
                   ladders_assigned = LADDERS_ASSIGNED_QUANTITY, others_units_assigned = OTHER_UNITS_ASSIGNED_QUANTITY)
```


As we can see from the summary there are many `NA` values, and many predictors that are as characters and not factors. In this step we will convert the characters predictors as factors merging the values that appear less in the dataset, so we do no have many values that have low frequency in our dataset.

In addition we will add he predictor for the `day_number`, a factorial predictor to indicate in the incident day is a week day or not `is_weekend` and a factorial predictor `time_of_day` that indicates the range of time whenever the incident happens, so `Night` (if the hour is between 0 and 6), `Morning`  (if the hour is between 6 and 12), `Afternoon` (if the hour is between 12 and 18), `Evening`  (if the hour is between 18 and 24).

Finally for convenience we convert the datetime that are currently a characters into POSIX type in the following format: `"%m-%d-%Y %H:%M:%S"`


```{r}
# set factorial
fire_data$inc_borough <- as.factor(fire_data$inc_borough)
fire_data$al_borough <- as.factor(fire_data$al_borough)
fire_data$al_source_desc <- as.factor(fire_data$al_source_desc)
fire_data$al_index_desc <- as.factor(fire_data$al_index_desc)
fire_data$highest_al_level <- as.factor(fire_data$highest_al_level)

fire_data$disp_resp_time_indc <- as.factor(fire_data$disp_resp_time_indc)
levels(fire_data$disp_resp_time_indc)<- c("N", "Y")

fire_data$inc_resp_sec_indc <- as.factor(fire_data$inc_resp_sec_indc)
levels(fire_data$inc_resp_sec_indc)<- c("N", "Y")

fire_data$inc_class_group <- as.factor(fire_data$inc_class_group)
fire_data$inc_class <- as.factor(fire_data$inc_class)
```

We note that the maximum level of the time indicator is very high to be considered as seconds so we decided to scale the two indicators in minutes.

```{r}
summary(fire_data %>% select(inc_resp_sec_qy, inc_travel_sec_qy))
```

```{r}
fire_data$inc_resp_sec_qy <- fire_data$inc_resp_sec_qy / 60
fire_data$inc_travel_sec_qy <- fire_data$inc_travel_sec_qy / 60
fire_data <- fire_data %>% rename(inc_resp_min_qy = inc_resp_sec_qy, inc_travel_min_qy = inc_travel_sec_qy)
```


Here we create the `day_number`, `time_of_day` and `is_weekend`


```{r}
#fire_data$datetime <- as.POSIXct(fire_data$datetime, format="%m/%d/%Y %H:%M:%S")
  
  # Process datetime column
fire_data$datetime <- mdy_hms(fire_data$datetime)
fire_data$inc_close_datetime <- mdy_hms(fire_data$inc_close_datetime)
  
# Create new columns for day number,  time of day and is weekend 
fire_data$day_number <- as.factor(day(fire_data$datetime))
  
fire_data$day_type <- as.factor(ifelse(weekdays(fire_data$datetime) %in% c("sabato", "domenica"), "Weekend", "Weekday"))
  
fire_data$ticket_time <- difftime(fire_data$inc_close_datetime, fire_data$datetime, units="mins")
  
fire_data$time_of_day <- cut(
    hour(fire_data$datetime),
    breaks = c(0, 6, 12, 18, 24),
    labels = c("Night", "Morning", "Afternoon", "Evening"),
    include.lowest = TRUE,
    right = TRUE
)
  
fire_data$datetime <- NULL
```


```{r}
table(fire_data$day_number)
```

```{r}
table(fire_data$time_of_day)
```

```{r}
table(fire_data$day_type)
```

Here we add a new predictor that is the sum of all the type of assigned units.

```{r}
fire_data$total_assigned_unit <- fire_data$engines_assigned + fire_data$ladders_assigned + fire_data$others_units_assigned
```


Set the dates to POSIX data type

```{r}
fire_data$first_onscene_datetime <- as.POSIXct(fire_data$first_onscene_datetime, format="%m/%d/%Y %H:%M:%S %p")
fire_data$first_ass_datetime <- as.POSIXct(fire_data$first_ass_datetime, format="%m/%d/%Y %H:%M:%S %p")
fire_data$first_act_datetime <- as.POSIXct(fire_data$first_act_datetime, format="%m/%d/%Y %H:%M:%S %p")
fire_data$inc_close_datetime <- as.POSIXct(fire_data$inc_close_datetime, format="%m/%d/%Y %H:%M:%S %p")
```


Rename the factor levels for the `inc_borough` predictor.

```{r}
fire_data <- fire_data %>% mutate(inc_borough = recode_factor(inc_borough, "BRONX" = "Bronx", "BROOKLYN" = "Brooklyn", "MANHATTAN" = "Manhattan", "QUEENS" = "Queens", "RICHMOND / STATEN ISLAND" = "Staten Island"))
```



At this point we merge some possible value from factorial predictors to make the space of possible choice smaller.

Here we merge the following factorial values of `highest_al_level`: `Second Alarm` and `Third Alarm` into `2nd-3rd Alarm`.


```{r}
# highest_al_level
fire_data$highest_alarm_lev_new <- fire_data$highest_al_level
levels(fire_data$highest_alarm_lev_new) <- list(
  "All Hands Working" = "All Hands Working",
  "First Alarm" = "First Alarm", 
  "2nd-3rd Alarm" = c("Second Alarm", "Third Alarm")
)

print(ctable(fire_data$highest_al_level, fire_data$highest_alarm_lev_new), method = 'render')

fire_data$highest_al_level <- fire_data$highest_alarm_lev_new
fire_data$highest_alarm_lev_new <- NULL
```

Here we merge the following factorial values of `al_index_desc`: `Second Alarm`, `Third Alarm`, `7-5 (All Hands Alarm)`, `10-76 & 10-77 Signal (Notification Hi-Rise Fire)` and `10-75 Signal (Request for all hands alarm)` into `Others`.

```{r}
# al_index_desc
fire_data$alarm_level_idx_new <- fire_data$al_index_desc
levels(fire_data$alarm_level_idx_new) <- list(
  "DEFAULT RECORD" = "DEFAULT RECORD",
  "Initial Alarm" = "Initial Alarm", 
  "Others" = c("Second Alarm", "Third Alarm", "7-5 (All Hands Alarm)", 
               "10-76 & 10-77 Signal (Notification Hi-Rise Fire)",
               "10-75 Signal (Request for all hands alarm)")
)

print(ctable(fire_data$al_index_desc, fire_data$alarm_level_idx_new), method = 'render')

fire_data$al_index_desc <- fire_data$alarm_level_idx_new
fire_data$alarm_level_idx_new <- NULL
```

Here we merge the following factorial values of `al_source_desc`: `911`, `911TEXT`, `VERBAL`, `BARS`, `ERS`, `ERS-NC` and `SOL` into `Others`.

```{r}
fire_data$alarm_source_desc_new <- fire_data$al_source_desc
levels(fire_data$alarm_source_desc_new) <- list(
  "PHONE" = "PHONE",
  "EMS" = "EMS",
  "EMS-911" = "EMS-911",
  "CLASS-3" = "CLASS-3",
  "Others" = c("911", "911TEXT", "VERBAL", "BARS", "ERS", "ERS-NC", "SOL")
)

print(ctable(fire_data$al_source_desc, fire_data$alarm_source_desc_new), method = 'render')

fire_data$al_source_desc <- fire_data$alarm_source_desc_new
fire_data$alarm_source_desc_new <- NULL
```
Visualise angain the dataset summary to see the changes.

```{r}
print(dfSummary(fire_data, 
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                graph.magnif = 0.76),
                method = 'render')
```



The next step is to deal with NA value and delete some un-useful predictors.


First of all we saw the possibility that `al_borough` and `inc_borough` are actually the same column.


```{r}
identical(fire_data$al_borough, fire_data$inc_borough)
```
The column ``al_borough` and `inc_borough` have the same sequence of values, so we can delete one of the two.

```{r}
fire_data <- fire_data %>% select(-c(al_borough))
```



Then we say that all observation in the dataset have the `disp_resp_time_indc` equal to *N*, let's check again and in affermative case we can delete both columns.

```{r}
summary(fire_data$disp_resp_time_indc)
```
All our observations have non valid `disp_resp_time_indc` so we can delete both the column indicator and the respective column quantity `inc_travel_min_qy`

```{r}
fire_data <- fire_data %>% select(-c(disp_resp_time_indc, inc_travel_min_qy))
```



Now we do a quick check also on the other indicator variable `inc_resp_sec_indc`

```{r}
summary(fire_data$inc_resp_sec_indc)
```
But here we have some observations with valid `inc_resp_sec_indc`, and we will consider only the valid one deleting the one that has a non valid attribute.

However before doing that let's be sure that the distribution of `inc_resp_min_qy` around the borough.

```{r}
ggplot(data=fire_data %>% group_by(inc_borough, inc_resp_sec_indc) %>% summarise(incident_number = n()), 
       aes(x=inc_borough, y=incident_number, fill=inc_resp_sec_indc)) +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_text(aes(label=incident_number), vjust=1.6, color="white",
            position = position_dodge(0.9), size=3.5) +
  scale_fill_brewer(palette="Paired") +
  labs(title = "Incident Count - Borouh - Valid Response Time in Minutes", x = "Borough", y = "Incident Number", fill = "Valid Response\n Time in Minutes") +
  theme_gray()
```

And to the rateo of valid `inc_resp_sec_indc` in each borough is:

```{r}
print(fire_data %>% group_by(inc_borough, inc_resp_sec_indc) %>% summarise(incident_number = n()) %>% mutate(ratio=incident_number/sum(incident_number))) %>% filter(inc_resp_sec_indc == "Y")
```



```{r}
ggplot(fire_data, aes(total_assigned_unit, inc_resp_min_qy)) + 
  geom_point(aes(colour = inc_resp_sec_indc))+
   labs(title = "Total Assigned Units - Response Time In Minutes", x = "Total Assigned Units", y = "Response Time In Minutes", colour = "Valid Response\n Time in Minutes") +
  theme_gray()
```

```{r}
ggplot(fire_data %>% filter(inc_resp_sec_indc == "N")
            , aes(total_assigned_unit, inc_resp_min_qy)) + 
  geom_point(aes(colour = inc_class_group)) +
  labs(title = "Total Assigned Units - Response Time In Minutes - Incidnet Class Group", x = "Total Assigned Units", y = "Response Time In Minutes", colour = "Incident Class Groups") +
  theme_gray()
```



```{r}
print(fire_data %>% 
        filter(inc_resp_sec_indc == "N", inc_class_group == "Medical Emergencies", total_assigned_unit == 1) %>%
        group_by(inc_borough) %>%
        summarise(incident_number = n()))

print(fire_data %>% 
        filter(inc_resp_sec_indc == "N", inc_class_group == "Medical Emergencies", total_assigned_unit != 1) %>%
        group_by(inc_borough) %>%
        summarise(incident_number = n()))

print(fire_data %>% 
        filter(inc_resp_sec_indc == "N", inc_class_group == "Medical Emergencies", total_assigned_unit == 1) %>%
        group_by(inc_class) %>%
        summarise(incident_number = n()))
```

```{r fig.height = 5, fig.width = 15}
ggplot(data=fire_data %>% 
        filter(inc_resp_sec_indc == "N", inc_class_group == "Medical Emergencies", total_assigned_unit == 1) %>%
        group_by(inc_class, inc_borough) %>%
        summarise(incident_number = n()), 
       aes(x=inc_borough, y=incident_number, fill=inc_class)) + geom_bar(stat="identity", position=position_dodge()) +
        geom_text(aes(label=incident_number), vjust=1.6, color="black",
                  position = position_dodge(0.9), size=3) +
        #scale_fill_brewer(palette="Paired") +
        labs(title = "Borough - Incident Counts - Incident Class Group", x = "Borough", y = "Incident Counts", fill = "Incident Class Group") +
        theme_grey()
```


We have concluded that the observations having `inc_resp_sec_indc == "N", inc_class_group == "Medical Emergencies", total_assigned_unit == 1` are mostly identified as **Medical - EMS Link 10-91** and **Medical - PD Link 10-91**.

1. **10-91 Medical Emergency EMS** - Fire Unit Not Required - To be transmitted through borough dispatcher by the responding unit when the fire Unit is canceled enroute due to EMS on scene, or EMS downgrades the job to a segment that does not require a Fire Unit response. Note: This signal shall be used only for medical emergency incidents. EMS stands for *Emergency Medical Services*.

2. **10-91 Medical Emergency PD** - Fire Unit Not Required - To be transmitted through borough dispatcher by the responding unit when the fire Unit is canceled enroute due to PD on scene, or PD downgrades the job to a segment that does not require a Fire Unit response. Note: This signal shall be used only for medical emergency incidents. PD stands for *Police Department*.


For and additional proof we can see also the relation with the `al_source_desc`.

```{r}
ggplot(fire_data %>% filter(inc_resp_sec_indc == "N")
            , aes(total_assigned_unit, inc_resp_min_qy)) +
  geom_point(aes(colour = factor(al_source_desc))) +
  labs(title = "Total Assigned Units - Response Time In Minutes - Alarm Source", x = "Total Assigned Units", y = "Response Time In Minutes", colour = "Alarm Source") +
  theme_gray()
```

And as expected the higher number of `inc_resp_min_qy` is from the `al_source_desc` equal to EMS and EMS-911. Whereas for the higher number of total assigned units and lower response time the majority of incident were reported by phone call.


Now we can look for the NonMedical Emergencies.



```{r}
print(fire_data %>% 
        filter(inc_resp_sec_indc == "N", inc_class_group == "NonMedical Emergencies") %>%
        group_by(inc_class) %>%
        summarise(incident_number = n()))
```
```{r}
ggplot(data=fire_data %>% 
          filter(inc_resp_sec_indc == "N", inc_class_group == "NonMedical Emergencies", inc_class == "Assist Civilian - Non-Medical") %>%
          group_by(inc_borough) %>%
          summarise(incident_number = n()), 
        aes(x=inc_borough, y=incident_number)) + 
      geom_bar(stat="identity", position=position_dodge()) +
      geom_text(aes(label=incident_number), vjust=1.6, color="white", position = position_dodge(0.9), size=3.5) +
      #scale_fill_brewer(palette="Paired") +
      labs(title = "Incident Count - Borouh - Valid Response Time in Second", x = "Borough", y = "Incident Count") +
      theme_minimal()
```

Where as the majority of non valid `inc_resp_sec_indc` that are Non-Medical Emergency are from the incident class equal to Assist Civilian - Non-Medical.



For stake of consistency we will consider only the valid observations that have `inc_resp_sec_indc == "Y"`.

```{r}
fire_data <- fire_data %>% filter(inc_resp_sec_indc == "Y")
dim(fire_data)
```


Now we want to know how many `incident_class` are summarized in each incident_class_group.

```{r}
print(ctable(fire_data$inc_class, fire_data$inc_class_group), method = 'render')
```

As we can see from the upper table all the `inc_class_group` have a unique set of values.

At this point to be more clear we display each main class with each respective sub-class.

```{r}
for (variable in levels(fire_data$inc_class_group)) {
  non_zero_table <- table(subset(fire_data, inc_class_group == variable)$inc_class)
  cat(variable, "\n")
  print(non_zero_table[non_zero_table != 0])
  cat("\n")
}
```


## NA Patterns?

At this point is essential to deal with NA values, trying to find the presence of possible relation with predictors.
First things first let's recap the number of NA values for each columns

```{r}
colSums(is.na(fire_data))
```

### Checking the location predictors

Here we will check if there is a pattern on the absence of values in the following predictors: `zipcode`, `pol_prec`, `city_con_dist`, `commu_dist`, `commu_sc_dist` and `cong_dist`.

```{r}
na_locations <- fire_data %>%
  filter(is.na(zipcode) | is.na(pol_prec) | is.na(city_con_dist) | is.na(commu_dist) | is.na(commu_sc_dist) | is.na(cong_dist))
```


```{r fig.height = 5, fig.width = 10}
ggplot(data=na_locations %>% 
        group_by(inc_class_group, inc_borough) %>%
        summarise(incident_number = n()), 
       aes(x=inc_borough, y=incident_number, fill=inc_class_group)) + geom_bar(stat="identity", position=position_dodge()) +
        geom_text(aes(label=incident_number), vjust=1.6, color="black",
                  position = position_dodge(0.9), size=3.5) +
        #scale_fill_brewer(palette="Paired") +
        labs(title = "NA location", x = "Borough", y = "Incident Count", fill = "Incident Class Group") +
        theme_grey()
```

```{r}
table(na_locations$inc_borough) / table(fire_data$inc_borough)
```

```{r}
table(na_locations$inc_class_group) / table(fire_data$inc_class_group)
```

So around the 40% of the whole incident that are of the incident class group NonMedical MFAs have at least one of the location columns to NA. Let's investigate.


```{r}
fd_nm_mfa_cl <- table(subset(fire_data, inc_class_group == "NonMedical MFAs")$inc_class)
fd_nm_mfa_bro <- table(subset(fire_data, inc_class_group == "NonMedical MFAs")$inc_borough)

fd_nm_mfa_cl <- fd_nm_mfa_cl[fd_nm_mfa_cl != 0]
fd_nm_mfa_cl
```


```{r}
na_nm_mfa_cl <- table(subset(na_locations, inc_class_group == "NonMedical MFAs")$inc_class)
na_nm_mfa_bro <- table(subset(na_locations, inc_class_group == "NonMedical MFAs")$inc_borough)

na_nm_mfa_cl <- na_nm_mfa_cl[names(fd_nm_mfa_cl)]
na_nm_mfa_cl
```



```{r}
na_nm_mfa_cl / fd_nm_mfa_cl
```
So the 97% of all the Non-Medical MFA - ERS observations in the entire dataset have one of the location attribute equal to NA

```{r}
na_nm_mfa_bro / fd_nm_mfa_bro
```
And from here we can see that about the 78% of the observations that are NonMedical - MFAs that have at least one district column attribute to NA are from the RICHMOND / STATEN ISLAND.
Also BRONX has about half of the NonMedical - MFAs observations that have at least one district column to NA.



### Checking the assigned units predictors

```{r}
print(fire_data %>%
  filter(is.na(engines_assigned) | is.na(ladders_assigned) | is.na(others_units_assigned)) %>%
  group_by(inc_borough, inc_class)) %>%
  summarise(incident_count = n())
```


### Checking the first_act_datetime predictors

```{r}
na_first_act_datetime <- fire_data %>% filter(is.na(first_act_datetime))
```

```{r}
print(na_first_act_datetime %>% group_by(inc_class, inc_borough) %>% summarise(incident_count = n()))
```


```{r}
ggplot(data=na_first_act_datetime %>% 
        group_by(inc_class_group, inc_borough) %>%
        summarise(incident_number = n()), 
       aes(x=inc_borough, y=incident_number, fill=inc_class_group)) + geom_bar(stat="identity", position=position_dodge()) +
labs(title = "NA First Act Date", x = "Borough", y = "Incident Count", fill = "Incident Class Group") +
  theme_minimal()

```

Seems to be random and thus there is no pattern that motivate the presence od na values in `first_act_datetime`.


At this point we can omit the na values.

```{r}
fire_data <- na.omit(fire_data)
```




## Additional Data Visaulization

In this section we plot additional data visualization focus on the geographical visualization of the New York borough with relative predictors. In order to do so we load two additional datasets:
1. **Alarm_Box_Locations.csv** is a dataset that includes geographical informations about the alarm box, including *latitude* and *longitude* useful to plot points in a map.
2. **fdny-firehouse-listing.csv** is a dataset that includes the geographical informations of every firefighter stations in the NYC, including again *latitude* and *longitude*.

```{r}
alarm_box_loc <- read.csv("datasets/Alarm_Box_Locations.csv")

head(alarm_box_loc)
```

```{r}
summary(alarm_box_loc)
```


```{r}
firefighter_stations <- read.csv("datasets/fdny-firehouse-listing.csv")

head(firefighter_stations)
```

```{r}
summary(firefighter_stations)
```


We now start with the firefighter stations dataset.

```{r}

# make a copy of the fire_data
fire_data_for_ffs <- fire_data

fire_data_for_ffs <- fire_data_for_ffs %>% rename(borough = inc_borough)

firefighter_stations$Borough <- as.factor(firefighter_stations$Borough)
firefighter_stations <- firefighter_stations %>% rename(borough = Borough)

# remove the na values
firefighter_stations <- na.omit(firefighter_stations)
```



```{r}
stations_borough <- firefighter_stations %>%
                    group_by(borough) %>%
                    summarise(number_of_stations = n())
```


```{r}
count_inc_brough <- fire_data_for_ffs %>% group_by(borough) %>% summarise(incident_count = n())

stations_borough$incident_per_station <- round(count_inc_brough$incident_count / stations_borough$number_of_stations, digits = 3)

count_inc_brough <- merge(count_inc_brough, stations_borough, by="borough")
```

```{r}
firefighter_stations <- firefighter_stations %>% rename(lat = Latitude, lon = Longitude)
firefighter_stations_sdf <- st_as_sf(firefighter_stations, coords = c("lon", "lat"), crs = 4326)
head(firefighter_stations_sdf)
```


Downloand of the GEO JSON

```{r}
geojson_newyork <- geojson_read("datasets/NYC_BoroughBoundaries.geojson",  what = "sp")
geojson_newyork <- setNames(geojson_newyork, c("borough_code", "borough", "shape_area", "shape_leng"))
geojson_newyork$borough <- as.factor(geojson_newyork$borough)
geojson_newyork$borough_code <- NULL
head(geojson_newyork)
```


```{r}
geojson_newyork@data = data.frame(geojson_newyork@data, count_inc_brough[match(geojson_newyork@data$borough, count_inc_brough$borough),])
geojson_newyork@data$borough.1 <- NULL
```


```{r}
mapview(list(firefighter_stations_sdf, geojson_newyork),
        zcol = list(NULL, "incident_count"),
        legend = list(FALSE, TRUE),
        homebutton = list(FALSE, TRUE), layer.name = list(NULL, "indicents_number"), alpha.regions = 0.5, aplha = 1)
```



distribution of incident per incident class
```{r fig.height = 5, fig.width = 10}
ggplot(data=fire_data %>% group_by(inc_class_group, inc_borough) %>%  summarise(incident_count = n()), 
       aes(x=inc_borough, y=incident_count, fill=inc_class_group)) +
    geom_bar(stat="identity", position=position_dodge()) +
    geom_text(aes(label=incident_count), vjust=1.6, color="black",
              position = position_dodge(0.9), size=2.5) +
    #scale_fill_brewer(palette="Paired") +
    labs(title = "Borough - Incident Count - Incident Class Group", x = "Borough", y = "Incident Count", fill = "Incident Class Group") +
  theme_minimal()
```



```{r fig.height = 5, fig.width = 8}
ggplot(fire_data, aes(x = inc_borough, y = inc_resp_min_qy, color = inc_class_group)) +  # ggplot function
  geom_boxplot() + scale_color_brewer(palette="Dark2") +
  labs(title = "Borough - Incident Class Group - Valid Response Time in Minutes", x = "Borough", y = "Valid Response Time in Minutes", color = "Incident Class Group") +
  theme_grey()
```



```{r}
fire_data[which.max(fire_data$inc_resp_min_qy), ]
```








```{r fig.height = 5, fig.width = 8}
ggplot(fire_data, aes(x = inc_borough, y = engines_assigned, color = inc_class_group)) +  # ggplot function
  geom_boxplot() +
  labs(title = "Borough - Engine Assigned - Incident Class Group", x = "Borough", y = "Engine Assigned", color = "Incident Class Group")
  
```

```{r}
fire_data %>% arrange(desc(engines_assigned)) %>% slice(1:2)
```

09/16/2023 06:59:35 PM, structural fires, source incident phone, url of the incident https://www.cbsnews.com/newyork/news/gravesend-brooklyn-fire-injuries/
09/24/2023 12:05:21 AM, structural fires, source incident phone, url of the incident https://abc7ny.com/three-injured-in-fire-bedford-stuyvesant-beford-stuyvesant-brooklyn-fdny/13823430/

```{r fig.height = 5, fig.width = 8}
ggplot(fire_data, aes(x = inc_borough, y = ladders_assigned, color = inc_class_group)) +  # ggplot function
  geom_boxplot() +
  labs(title = "Borough - Ladders Assigned - Incident Class Group", x = "Borough", y = "Engine Assigned", color = "Incident Class Group")
```

```{r}
fire_data %>% arrange(desc(ladders_assigned)) %>% slice(1:2)
```

Same incidents of before

```{r fig.height = 5, fig.width = 8}
ggplot(fire_data, aes(x = inc_borough, y = others_units_assigned, color = inc_class_group)) +  # ggplot function
  geom_boxplot() +
  labs(title = "Borough - Other Units Assigned - Incident Class Group", x = "Borough", y = "Other Units Assigned", color = "Incident Class Group")
```

```{r}
fire_data %>% arrange(desc(others_units_assigned)) %>% slice(1:2)
```

Same incidents of before



```{r fig.height = 5, fig.width = 8}
ggplot(fire_data, aes(x = inc_borough, y = total_assigned_unit, color = inc_class_group)) +  # ggplot function
  geom_boxplot() + 
  labs(title = "Borough - Total Units Assigned - Incident Class Group", x = "Borough", y = "Total Units Assigned", color = "Incident Class Group")
```

```{r}
fire_data %>% arrange(desc(total_assigned_unit)) %>% slice(1:2)
```

Same incidents of before



```{r}
ggplot(fire_data %>% group_by(al_source_desc, inc_class_group, inc_borough) %>% summarise(incidents_number = n()),
       aes(x = inc_borough, y = incidents_number, color = al_source_desc)) +  # ggplot function
  geom_boxplot() +
  labs(title = "Borough - Incident Number - Alarm Source", x = "Borough", y = "Incident Number", color = "Alarm Source")
```

```{r}
ggplot(fire_data %>% group_by(inc_class_group, inc_borough, day_type) %>% summarise(incidents_number = n()),
       aes(x = inc_borough, y = incidents_number, color = day_type)) +
  geom_boxplot()  +
  labs(title = "Borouhg - Incident Count - Day Type", x = "Borough", y = "Incident Count", color = "Day Type")
```


```{r}
ggplot(fire_data %>% group_by(inc_class_group, inc_borough, time_of_day) %>% summarise(incidents_number = n()),
       aes(x = inc_borough, y = incidents_number, color = time_of_day)) +
  geom_boxplot() +
  labs(title = "Borough - Incident Count - Day Time", x = "Borough", y = "Incident Count", color = "Day Time")
```



# Let's build some models

But before creating any model have to split the cleaned dataset into *train* and *test*, with 0.8% of the whole dataset for the train set and the remaining 20% for the test set.

```{r}
sample <- sample(c(TRUE, FALSE), nrow(fire_data), replace=TRUE, prob=c(0.8,0.2))
fire_data.train <- fire_data[sample, ]
fire_data.test <- fire_data[!sample, ]
dim(fire_data.train)
dim(fire_data.test)
```


# Molutinomial Logistic Regression
```{r}
pValue_extract <- function(x){
  z <- summary(x)$coefficients/summary(x)$standard.errors
  # 2-tailed Wald z tests to test significance of coefficients
  p <- (1 - pnorm(abs(z), 0, 1)) * 2
  p
}
```


```{r}
multi.fit1 <- multinom(incident_class ~ ., data = fire_data.train, trace = FALSE)
```

```{r}
table(preds = multi.preds1 <- predict(multi.fit1, newdata = fire_data.test), true = fire_data.test$incident_class)
```

```{r}
multi.fit1$AIC
```
```{r}
mean(multi.preds1 == fire_data.test$incident_class)
```

```{r}
#pValue_extract(multi.fit1)
```




Try a new model with less predictors
```{r}
multi.fit2 <- update(multi.fit1, . ~ . - AL_loction - zip - Pprecint - citycouncil_dist - community_dist - comm_school_dist - congressional_dist - day_number)
multi.fit2$AIC
```

```{r}
table(preds = multi.preds2 <- predict(multi.fit2, newdata = fire_data.test), true = fire_data.test$incident_class)
```


```{r}
mean(multi.preds2 == fire_data.test$incident_class)
```

```{r}
summary(multi.fit2, wald = TRUE)
```

```{r}
#pValue_extract(multi.fit2)
```


# Linear Discriminant Analysis
Now consider linear discriminant analysis:
```{r}
library("MASS")
lda.fit <- lda(formula(multi.fit2), data = fire_data.train)
lda.fit
```
Prediction:
```{r}
lda.preds <- predict(lda.fit, newdata = fire_data.test)
``` 

```{r}
table(preds = lda.preds$class, true = fire_data.test$incident_class)
```
The accuracy of linear discriminant analysis is worse respect to the multinomial logistic regression:
```{r}
mean(lda.preds$class == fire_data.test$incident_class)
```


Now consider quadratic discriminant analysis:



# Naive Bayes
```{r}
library("e1071")
nb.fit <- naiveBayes(formula(multi.fit2), data = fire_data.train)
```
Prediction:
```{r}
nb.preds <- predict(nb.fit, newdata = fire_data.test, type = "class")
``` 

```{r}
table(preds = nb.preds, true = fire_data.test$incident_class)
```
The accuracy of linear discriminant analysis is worse respect to the multinomial logistic regression:
```{r}
mean(nb.preds == fire_data.test$incident_class)
```

# K-Nearest Neighbours
```{r}
x.train <- model.matrix(~ incident_brough + alarm_source_desc + alarm_level_idx + 
    highest_alarm_lev + incident_resp_sec + incident_travel_time_sec + 
    engines_assigned + ladders_assigned + other_utits_assigned + 
    time_of_day, data = fire_data.train)[, -1]
x.test <- model.matrix(~ incident_brough + alarm_source_desc + alarm_level_idx + 
    highest_alarm_lev + incident_resp_sec + incident_travel_time_sec + 
    engines_assigned + ladders_assigned + other_utits_assigned + 
    time_of_day, data = fire_data.test)[, -1]
library("class")
set.seed(98765)
rates <- double(30)
for (i in 1:30) {
  tmp <- knn(train = x.train, test = x.test, cl = fire_data.train$incident_class, k = i)
  rates[i] <- mean(tmp == fire_data.test$incident_class)
}
plot(x = (1:30), y = rates, xlab = "k", ylab = "Accuracy", type = "l")
```


